A weighted sampling algorithm to designing RNA sequences with target secondary structure and nucleotide distribution

### 2013-01-08 ISMB/results_ISMB.tex
f

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\textwidth]{Figures/time_rnastrand_clustered_rnainverse_100samples_fix}
	\caption{Average time in seconds to generate one sequence. We explicitly show 
	the time spent by \ourprog (full line) and \texttt{RNAinverse} (dashed) for various \GCContent. The first plot is in function
	of the length of the structures. The second is in function of the stack
	density (i.e. $2\cdot\#stacks/length$) and the latter in function of 
	the proportion of free nucleotides (i.e. not base paired) inside loops.}
	\label{fig:time}	
\end{figure}

\subsection{Dataset}
To evaluate the quality of our method, we used secondary
structures from the \RNASTRAND database~\cite{andronescu2008rna}.
Those are known secondary structure from a variety of organisms.
We select the same subset of $50$ structures as in~\cite{Levin:2012kx}, 
which have lengths between $20$ and $100$ nucleotides.

 To ease the visualization of the data, we clustered together the strucures
 in function of their length, stacks density and proportion of free 
 nucleotides in loops. We present in Figure~\ref{fig:bins} the distribution
 of the structures in the different bins.

 \begin{figure}[ht!]
 	\centering
	\includegraphics[width=\textwidth]{Figures/bins_distribution.png}
	\caption{Number of secondary structures per bins.}
	\label{fig:bins}
 \end{figure}

\subsection{Design}
 To benchmark our method, we first started by sampling a minimum
 of $100$ sequence per structure. After correcting the free nucleotides with
 \texttt{RNAinverse} we computed their MFE with \texttt{RNAfold} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.

A first criterion is which percentage of sequences have as MFE the target
secondary structures. We are also interested by the number of structures
for which we can design at least one sequence with the desired MFE.
We show the results when using only \ourprog in 
 Figure~\ref{fig:mfe_struct_solved_noinverse}. As mentioned earlier, the high rate of failure
is due to the fact that no selection criterion is applied to
unpaired nucleotides. A local strategy is thus needed.
After processing \ourprog sequences with \RNAinverse to 
locally optimize the free nucleotides, we obtained the results 
shown in Figure~\ref{fig:mfe_struct_solved}. We observed
that length is not a good predictor for the quality of the results. Instead,
a high number of free nucleotides in the structure seems to be a 
good measure of its design hardness. 
 The difficulty of 
designing sequences for targets with a high number of free nucleotides 
 in loops appears clearly in the last column of Figure~\ref{fig:mfe_struct_solved}.
Even with a \GCContent of $50\%$ and more, when at least
one solutions was found for almost all structures, most of the samples 
do not fold into the target. A few structures remained unsolved under
all \GCContent.

@@

A desirable feature in sequence design, is to produce samples with a high
diversity. We show in Figure~\ref{fig:nb_sols_entropy} the number of correct
solutions as their average entropy and base pair entropy, since 
\ourprog  only influences the distribution of nucleotides inside 
stacked base pairs. The most constraining factor for generating valid
 solutions seems to be  low \GCContent and the percentage of free nucleotides. Our method is able to generate

@@

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\textwidth]{Figures/freq.png}
	\caption{Average frequency of the MFE.}
	\label{fig:freq}
\end{figure}

\subsection{Global sampling vs Local search vs Glocal approach}
Here, we are interested in estimating the impact of the design methodology on the performances. More precisely, we aim to what are the benefits of a global sampling approach (\ourprog), versus a glocal procedure (\ourprog + \RNAinverse) or the classical local search methodology (\RNAinverse). Thus, we compare our method to a pure local search procedure. In particular, here we compare to \RNAinverse as it is the program we used to do the local optimization (see Section \ref{subsec:glocal_method}). Our results (See Fig. \ref{fig:rnainverse} in Supplementary data) show that \ourprog overcome other strategies. In particular, the base pair entropy as well as diversity of sequences are favorable to \ourprog. This advantage is even more striking for low \GCContent regimes.

To complete this benchmark, we added the results obtained with the glocal procedure (\ourprog + \RNAinverse). We note that the results are slightly lower than those of \ourprog alone. However, we must stress that the \ourprog data consider only the sample of \ourprog already satisfying the MFE criterion. In fact \RNAinverse enable to ``correct'' a lot of sample sampled by \ourprog but without the good folding properties. Hence, our conclusion here is that the glocal approach \ourprog + \RNAinverse enable us to conserve the same level of performance but to drastically improve the success rate of our methodology.

>>>>

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\textwidth]{Figures/time_rnastrand_clustered_rnainverse_100samples_fix}
	\caption{Average time in seconds to generate one sequence. We explicitly show 
	the time spent by \ourprog (full line) and \texttt{RNAinverse} (dashed) for various \GCContent. The first plot is in function
	of the length of the structures. The second is in function of the stack
	density (i.e. $2\cdot\#stacks/length$) and the latter in function of 
	the proportion of free nucleotides (i.e. unpaired bases) within loops.}
	\label{fig:time}	
\end{figure}

\subsection{Dataset}
To evaluate the quality of our method, we used secondary
structures from the \RNASTRAND database~\cite{andronescu2008rna}.
Those are known secondary structure from a variety of organisms.
We select the same subset of $50$ structures as in~\cite{Levin:2012kx}, 
whose length ranges between $20$ and $100$ nucleotides.

 To ease the visualization of the results, we clustered together strucures
 having similar length, stacks density and proportion of free-nucleotides in loops. 
  We present in Figure~\ref{fig:bins} the distribution of structures in each category.

 \begin{figure}[ht!]
 	\centering
	\includegraphics[width=\textwidth]{Figures/bins_distribution.png}
	\caption{Number of secondary structures per bin, according to our three clustering criteria.}
	\label{fig:bins}
 \end{figure}

\subsection{Design}
 To benchmark our method, we first started by sampling about $100$ sequences per structure. After correcting the free nucleotides with
 \texttt{RNAinverse} we computed their MFE with \texttt{RNAfold} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.

A first criterion is which percentage of sequences have as MFE the target
secondary structures. We are also interested by the number of structures
for which we can design at least one sequence with the desired MFE.
We show the results when using only \ourprog in 
 Figure~\ref{fig:mfe_struct_solved_noinverse}. As mentioned earlier, the high rate of failure
is due to the fact that no selection criterion is applied to
unpaired nucleotides. A local strategy is thus needed.
After processing \ourprog sequences with \RNAinverse to 
locally optimize the free nucleotides, we obtained the results 
shown in Figure~\ref{fig:mfe_struct_solved}. We observed
that length is, in general, not a good predictor for the hardness of designing a structure. 
Instead, a high number of free nucleotides in the structure seems to be a 
good measure of its design hardness. 
 The difficulty of 
designing sequences for targets with a high number of free nucleotides 
 in loops appears clearly in the last column of Figure~\ref{fig:mfe_struct_solved}.
Even with a \GCContent of $50\%$ and more, when at least
one solutions was found for almost all structures, most of the samples 
do not fold into the target. A few structures remained unsolved under
all \GCContent.

@@

A desirable feature in sequence design, is to produce samples with a high
diversity. Figure~\ref{fig:nb_sols_entropy} shows the number of correct
solutions as their average entropy and base pair entropy, since 
\ourprog  only influences the distribution of nucleotides inside 
stacked base pairs. The most constraining factor for generating valid
 solutions seems to be  low \GCContent and the percentage of free nucleotides.

%Our method is able to generate

@@

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\textwidth]{Figures/freq.png}
	\caption{Average Boltzmann probability of the MFE.}
	\label{fig:freq}
\end{figure}

\subsection{Global sampling vs Local search vs Glocal approach}

Here, we are interested in estimating the impact of the design methodology on the performances. More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and a local search methodology (\RNAinverse).

Our results (See Fig. \ref{fig:rnainverse} in Supplementary Material) indicate that a glocal approach produces sequences that are consistently more diverse than those obtained with both {\em pure} methods alone. Furthermore, while \ourprog outperforms \RNAinverse for extreme pressures on the \GCContent, \RNAinverse produces more entropic sequences than \ourprog for the central \GCContent regime ($30\%-70\%$).

It should be noted, however, than \RNAinverse produces sequences whose \GCContent is extremely concentrated (see Table~\ref{tab:nb_rnainv} in Supplementary Material), thereby requiring a large number of runs before producing a suitable \GCContent sequence. Therefore, a rejection algorithm based on \RNAinverse alone should not be considered a realistic solution for the design with constrained \GCContent.

% \ourprog outperforms other strategies for non-centered . In particular, the base pair entropy as well as diversity of sequences are favorable to \ourprog. This advantage is even more striking for low targeted \GCContent{}s regimes.
%To complete this benchmark, we added the results obtained with the glocal procedure (\ourprog + \RNAinverse). We note that the results are slightly lower than those of \ourprog alone. However, we must stress that the \ourprog data consider only the sample of \ourprog already satisfying the MFE criterion. In fact \RNAinverse enable to ``correct'' a lot of sample sampled by \ourprog but without the good folding properties. Hence, our conclusion here is that the glocal approach enable us to conserve the same level of performance but to drastically improve the success rate of our methodology.
